JD notes on effects

### 2022-10-05 manuscript/paper.tex
Merge with Bicko?

\fref{splash} shows effect- and prediction-style confidence intervals for a hypothetical data set. Prediction-style confidence intervals do not depend on an anchor choice, and correspond to \CIs\ for the predicted mean value of Biomass for the given value of Nitrogen (note that these differ the broader \emph{prediction intervals} which give \CIs\ for the observations rather than the predicted means). Effect-style confidence intervals show the range of slopes corresponding to the model's \CIs\ for the \emph{direct} effect of the focal predictor.

@@

\begin{figure}
\begin{center}
\end{center}
\{fig:justify_plots}
\end{figure}
%
\fref{center} shows an example, based on simulated data with an interaction between correlated variables.
In the absence of interactions (\EREF{simple_inter_higher_no_interaction}), the three approaches produce identical estimates, which match the simulated values, \FREF{justify_plots}A. However, the estimates start to differ in the presence of interactions, even as simple as the one in \EREF{simple_inter_higher}. In particular, estimates from \pkg{emmeans} and \pkg{effects} are identical but differ from \pkg{varpred}'s, which is very close to the simulated average ($\bar{y}$), \FREF{justify_plots}B.
In the simple model, \FREF{justify_plots}A, the input variables are the same as the linear predictor variables, so all the three methods produce identical results.
In the interaction model, \FREF{justify_plots}B, there is an additional linear predictor variable ($x_2x_3$). \pkg{emmeans} and \pkg{effects} first average the input variables to compute $\bar{x_2}\bar{x_3}$ while \pkg{varpred} first calculates the corresponding vector of linear predictor values and then averages.

@@

\begin{figure}
\begin{center}
\includegraphics{eco.mcomp.Rout.pdf}
\end{center}
\caption{Outcome plots with effect- (dashed straight lines) and prediction-style (dashed curved lines). The panel on the left shows the results of a univariate model fit; the panel on the right shows a multivariate fit (and effect-style lines only, see text.}
\flab{mcomp}
\end{figure}
%%
\jd{Alternative figure 1?}
\fref{mcomp} shows effect- and prediction-style confidence intervals for a hypothetical data set. Prediction-style confidence intervals do not depend on an anchor choice, and correspond to \CIs\ for the predicted mean value of Biomass for the given value of Nitrogen (note that these differ the broader \emph{prediction intervals} which give \CIs\ for the observations rather than the predicted means). Effect-style confidence intervals show the range of slopes corresponding to the model's \CIs\ for the \emph{direct} effect of the focal predictor.

>>>>

\fref{splash} shows effect- and prediction-style confidence intervals for a hypothetical data set. Prediction-style confidence intervals do not depend on an anchor choice, and correspond to \CIs\ for the predicted mean value of Biomass for the given value of Nitrogen (note that these differ from the broader \emph{prediction intervals} which give \CIs\ for observations rather than the predicted means, and are not discussed further here). Effect-style confidence intervals show the range of slopes corresponding to the model's \CIs\ for the \emph{direct} effect of the focal predictor. In this simple case, effect-style CIs, unlike prediction-style CIs, have a clear visual correspondence with statistical clarity: both CI lines have the same slope exactly when the P value for the effect of the predictor is below the threshold.

@@

\begin{figure}
\begin{center}
\end{center}
\flab{justify_plots}
\end{figure}
%
\fref{center} shows an example, based on simulated data with an interaction between correlated variables.
In the absence of interactions (\EREF{simple_inter_higher_no_interaction}), the three approaches produce identical estimates, which match the simulated values, \FREF{justify_plots}A. However, the estimates start to differ in the presence of interactions, even as simple as the one in \EREF{simple_inter_higher}. In particular, estimates from \pkg{emmeans} and \pkg{effects} are identical but differ from \pkg{varpred}'s, which is very close to the simulated average ($\bar{y}$), \FREF{justify_plots}B.
In the simple model, \FREF{justify_plots}A, the input variables are the same as the linear predictor variables, so all the three methods produce identical results.
In the interaction model, \FREF{justify_plots}B, there is an additional linear predictor variable ($x_2x_3$). \pkg{emmeans} and \pkg{effects} first average the input variables to compute $\bar{x_2}\bar{x_3}$ while \pkg{varpred} first calculates the corresponding vector of linear predictor values and then averages.

@@

\begin{figure}
\begin{center}
\includegraphics{eco.mcomp.Rout.pdf}
\end{center}
\caption{Outcome plots with effect- (dashed straight lines) and prediction-style (dashed curved lines). The panel on the left shows the results of a univariate model fit; the panel on the right shows a multivariate fit (and effect-style lines only, see text.}
\flab{mcomp}
\end{figure}
%%
\jd{Still working on this. One panel or two?}
\fref{mcomp} compares direct and indirect effects using univariate vs.~multivariate models
